---
title: "Basic Inferential Data Analysis"
author: "Will Swales"
date: "10 July 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
The purpose of this report is to document an exploratory analysis of the ToothGrowth dataset.

## Raw Data
The raw data is the ToothGrowth dataset provided in the R datasets package. This is a data frame which contains data of an experiment to explore the effectof Vitamin C on Tooth Growth in guinea pigs.

The dataset comes with the following description:

>*The response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, orange juice or ascorbic acid (a form of vitamin C and coded as VC).*

## Data Summary
### Data Structure
```{r, echo=FALSE}
data("ToothGrowth")
str(ToothGrowth)
```

### Data Visualisation
```{r, echo=FALSE, warning=FALSE}
library(ggplot2)
qplot(dose,len,data=ToothGrowth,color=supp, geom=c("point","smooth"), method="lm", facets=.~supp)
```
